---
title: "Exploring Red Wine Quality"
author: "Shane Kao"
date: "Wednesday, February 18, 2015"
output: pdf_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(GGally)
library(corrplot)
library(scales)
library(memisc)
library(reshape)
library(gridExtra)
library(fmsb)
```

# Goal
Which chemical properties influence the quality of red wines?

# Data Overview
This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).
```{r}
setwd("C:/Users/asus/Downloads")
data=read.csv("wineQualityReds.csv",stringsAsFactors=FALSE)
str(data)
summary(data)
```

* The feature `X` is row index of data, it may provide no further information.
* The feature `quality` is an ordered, categorical, discrete variable.
* From the variable descriptions, {`fixed.acidity`,`volatile.acidity`,`citric.acid`} and {`free.sulfur.dioxide`,`total.sulfur.dioxide`} may strongly correlated.

# Univariate Plots Section

## Univariate Analysis 

The following figures shows distributions for all features:
```{r , message=FALSE, warning=FALSE,echo=FALSE}
grid.arrange(qplot(data$fixed.acidity,xlab="fixed acidity"),
             qplot(data$volatile.acidity,xlab="volatile acidity"),
             qplot(data$citric.acid,xlab="citric acid"),
             qplot(data$residual.sugar,xlab="residual sugar"),
             qplot(data$chlorides,xlab="chlorides"),
             qplot(data$free.sulfur.dioxide,xlab="free sulfur dioxide"),
             qplot(data$total.sulfur.dioxide,xlab="total sulfur dioxide"),
             qplot(data$density,xlab="density"),
             qplot(data$pH,xlab="pH"),
             qplot(data$sulphates,xlab="sulphates"),
             qplot(data$alcohol,xlab="alcohol"),
             qplot(data$quality,xlab="quality"),
             ncol = 3)
```

Now, we display density estimation for each features

```{r , message=FALSE, warning=FALSE,echo=FALSE}
                qplot(data$fixed.acidity,
                   xlab="fixed acidity", geom="density")
```

The sample average of `fixed.acidity` is about 8 

```{r , message=FALSE, warning=FALSE,echo=FALSE}
                qplot(data$volatile.acidity,
                   xlab="volatile acidity", geom="density")
```

The distribution of `volatile.acidity` reveal bimodal

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$citric.acid,
                   xlab="citric acid", geom="density")
```

The distribution of `citric.acid` reveal trimodal

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$residual.sugar,
                   xlab="residual sugar", geom="density")
```

Most of the `residual.sugar` is smaller than 4

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$chlorides,
                   xlab="chlorides", geom="density")
```

Most of the `chlorides` is smaller than 0.1

```{r , message=FALSE, warning=FALSE,echo=FALSE}
             qplot(data$free.sulfur.dioxide,
                   xlab="free sulfur dioxide", geom="density")
```

The mode of `free.sulfur.dioxide` is about 5

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$total.sulfur.dioxide,
                   xlab="total sulfur dioxide", geom="density")
```

The shape of distribution for `total.sulfur.dioxide` and `free.sulfur.dioxide` are similar

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$density,
                   xlab="wine density", geom="density")
```

`density` appear to be normally-distributed.

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$pH,
                   xlab="pH", geom="density")
```

* `pH` appear to be normally-distributed.

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$sulphates,
                   xlab="sulphates", geom="density")
```

The mode of `sulphates` is about 0.6

```{r , message=FALSE, warning=FALSE,echo=FALSE}
             qplot(data$alcohol,
                   xlab="alcohol", geom="density")
```

The mode of `alcohol` is about 9

```{r , message=FALSE, warning=FALSE,echo=FALSE}             
             qplot(data$quality,
                   xlab="quality", geom="density")
```

rating 5 and rating 6  has more points than other rating. 

## Short questions

### Did you create any new variables from existing variables in the dataset?
* It is convenient to interpret the result by creating variable `rating` , classifying each wine as low, medium and high, assign `quality` 3 and 4 to low level, 5 and 6 to medium and 7 and 8 to high.

```{r}
data$rating<-rep("",dim(data)[1])
data$rating[which(data$quality%in%3:4)]<-"low"
data$rating[which(data$quality%in%5:6)]<-"medium"
data$rating[which(data$quality%in%7:8)]<-"high"
table(data$rating)
```

*  I create a combined variable, `acid`, taking average of `fixed.acidity`, `volatile.acidity` and `citric.acid` after standardization. 

```{r}
data$acid<-((data$fixed.acidity-mean(data$fixed.acidity))/
                    sd(data$fixed.acidity)+
            (data$volatile.acidity-mean(data$volatile.acidity))/
                    sd(data$volatile.acidity)+
            (data$citric.acid-mean(data$citric.acid))/
                    sd(data$citric.acid))/3
```

*  I create a combined variable, `dioxide`, taking average of `free.sulfur.dioxide` and `total.sulfur.dioxide` after standardization. 

```{r}
data$dioxide<-((data$free.sulfur.dioxide-mean(data$free.sulfur.dioxide))/
                       sd(data$free.sulfur.dioxide)+
               (data$total.sulfur.dioxide-mean(data$total.sulfur.dioxide))/
                       sd(data$total.sulfur.dioxide))/2
```

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

```{r , message=FALSE, warning=FALSE, fig.height=20,,echo=FALSE}
grid.arrange(qplot(x = factor(""),y=data$fixed.acidity,geom='boxplot',
                   ylab="fixed acidity",xlab=""),
             qplot(x = factor(""),y=data$volatile.acidity,geom='boxplot',
                   ylab="volatile acidity",xlab=""),
             qplot(x = factor(""),y=data$citric.acid,geom='boxplot',
                   ylab="citric acid",xlab=""),
             qplot(x = factor(""),y=data$residual.sugar,geom='boxplot',
                   ylab="residual sugar",xlab=""),
             qplot(x = factor(""),y=data$chlorides,geom='boxplot',
                   ylab="chlorides",xlab=""),
             qplot(x = factor(""),y=data$free.sulfur.dioxide,geom='boxplot',
                   ylab="free sulfur dioxide",xlab=""),
             qplot(x = factor(""),y=data$total.sulfur.dioxide,geom='boxplot',
                   ylab="total sulfur dioxide",xlab=""),
             qplot(x = factor(""),y=data$density,geom='boxplot',
                   ylab="density",xlab=""),
             qplot(x = factor(""),y=data$pH,geom='boxplot',
                   ylab="pH",xlab=""),
             qplot(x = factor(""),y=data$sulphates,geom='boxplot',
                   ylab="sulphates",xlab=""),
             qplot(x = factor(""),y=data$alcohol,geom='boxplot',
                   ylab="alcohol",xlab=""),
             qplot(x = factor(""),y=data$quality,geom='boxplot',
                   ylab="quality",xlab=""),
             ncol = 3)

```

* `residual.sugar` and `chlorides` have extreme outliers.
* I don't tidy or adjust any data.

# Bivariate Plots Section

## Bivariate Analysis

We use the following figure to check the correlation between different features

```{r , message=FALSE, warning=FALSE,echo=FALSE}
cor_jnk=cor(data[,-c(1,2,3,4,7,8,13,14)], use="complete.obs")
corrplot(cor_jnk, order="AOE", method="circle", tl.pos="lt", type="upper",        
         tl.col="black", tl.cex=0.6, tl.srt=45, 
         addCoef.col="black", addCoefasPercent = TRUE,
         p.mat = 1-abs(cor_jnk), sig.level=0.50, insig = "blank")
```

Most of features are un-correlated, but `acid` is correlated with `pH` and `density`. Note that there are only two points colored off the main
diagonal (acid vs. density and acid vs ph) because corresponding correlation coefficient is regarded as significant.

The following boxplot shows how the features affect `rating`:

```{r , message=FALSE, warning=FALSE, fig.height=20,echo=FALSE}
grid.arrange(qplot(x = data$rating,y=data$acid,geom='boxplot',
                   ylab="acid",xlab="rating"),
             qplot(x = data$rating,y=data$residual.sugar,geom='boxplot',
                   ylab="residual sugar",xlab="rating"),
             qplot(x = data$rating,y=data$chlorides,geom='boxplot',
                   ylab="chlorides",xlab="rating"),
             qplot(x = data$rating,y=data$dioxide,geom='boxplot',
                   ylab="dioxide",xlab="rating"),
             qplot(x = data$rating,y=data$density,geom='boxplot',
                   ylab="density",xlab="rating"),
             qplot(x = data$rating,y=data$pH,geom='boxplot',
                   ylab="pH",xlab="rating"),
             qplot(x = data$rating,y=data$sulphates,geom='boxplot',
                   ylab="sulphates",xlab="rating"),
             qplot(x = data$rating,y=data$alcohol,geom='boxplot',
                   ylab="alcohol",xlab="rating"),
             ncol = 3)
```

* Higher `alcohol`, higher `rating`.
* Higher `sulphates`, higher `rating`.
* Lower `pH`, higher `rating`.
* Lower `density`, higher `rating`.

We use `quality` to support above result

```{r,echo=FALSE}
ggplot(data, aes(x =quality, y = alcohol)) +
        geom_point(position = 'jitter') +    
        geom_smooth(method=lm)       
ggplot(data, aes(x =quality, y = sulphates)) +
        geom_point(position = 'jitter')+
        geom_smooth(method=lm)               
ggplot(data, aes(x =quality, y = pH)) +
        geom_point(position = 'jitter')+
        geom_smooth(method=lm) 
ggplot(data, aes(x =quality, y = density)) +
        geom_point(position = 'jitter')+
        geom_smooth(method=lm) 
```

After using `quality`, we can see `alcohol` affect `quality` strongest. 

* Higher `alcohol`, higher `rating`.
* Higher `sulphates`, higher `rating`.
* Lower `pH`, higher `rating`.
* Lower `density`, higher `rating`.

In Univariate Plots Section, we conclude that `density` and `pH` appear to be normally-distributed. Now we use the following plots to support the result

```{r echo=FALSE}
ggplot(data, aes(density, colour = rating)) +
        geom_density()+xlab("Wine density")
ggplot(data, aes(pH, colour = rating)) +
        geom_density()
```

It seems like wine with medium level appear to normally-distributed. The result is not suprised because medium level have larger sample size.

## Short questions

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

We find out `alcohol`, `sulphates`, `density` and `pH` affect `rating` or `quality`.

### What was the strongest relationship you found?

Higher `alcohol`, higher `rating`.

# Multivariate Plots Section

## Multivariate Analysis

I examined the scatter plot of all the pair of features containing `alcohol`, `sulphates`, `density` and `pH` between different `quality`.

```{r ,message=FALSE,warning=FALSE,echo=FALSE}
ggplot(data = data,aes(x = alcohol, y = sulphates))+
       geom_point(aes(color =quality))+
       scale_color_continuous(name="quality",low = "white",high = "red")  
```

Lower `sulphates` and lower `alcohol`, lower `quality`.

```{r ,message=FALSE,warning=FALSE,echo=FALSE}
ggplot(data = data,aes(x = alcohol, y = density))+
       geom_point(aes(color =quality))+
       scale_color_continuous(name="quality",low = "white",high = "red") 
```

Higher `density` and lower `alcohol`, lower `quality`.

```{r ,message=FALSE,warning=FALSE,echo=FALSE}
ggplot(data = data,aes(x = alcohol, y = pH))+
       geom_point(aes(color =quality))+
       scale_color_continuous(name="quality",low = "white",high = "red") 
```

Higher `pH` and lower `alcohol`, lower `quality`.

```{r ,message=FALSE,warning=FALSE,echo=FALSE}
ggplot(data = data,aes(x = sulphates, y = density))+
       geom_point(aes(color =quality))+
       scale_color_continuous(name="quality",low = "white",high = "red") 
```

Lower `sulphates`, lower `quality`.

```{r ,message=FALSE,warning=FALSE,echo=FALSE}
ggplot(data = data,aes(x = sulphates, y = pH))+
       geom_point(aes(color =quality))+
       scale_color_continuous(name="quality",low = "white",high = "red") 
```

Lower `sulphates`, lower `quality`.


```{r ,message=FALSE,warning=FALSE,echo=FALSE}

ggplot(data = data,aes(x = density, y = pH))+
       geom_point(aes(color =quality))+
       scale_color_continuous(name="quality",low = "white",high = "red") 
```

Higher `pH` and higher `density`, lower `quality`.


The above plots support the result in Bivariate Plots Section.

## Short questions

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

Higher `density` and lower `alcohol`, lower `quality`.

# Final Plots and Summary

## Plot One
```{r echo=FALSE,message=FALSE,warning=FALSE, Plot_One}
grid.arrange(qplot(data$density,xlab="density",main='distribution of density'),
             qplot(data$pH,xlab="pH",main='distribution of pH'),
             ncol = 2)
```

## Description One

`density` and `pH` appear to be normally-distributed.

### Plot Two
```{r echo=FALSE,message=FALSE,warning=FALSE, Plot_Two}
ggplot(data, aes(x =quality, y = alcohol)) +
       geom_point(position = 'jitter') +    
       geom_smooth(method=lm)+
       ggtitle("alcohol vs quality and regression line")
```

### Description Two

Higher `alcohol`, higher `quality`.

### Plot Three
```{r echo=FALSE,message=FALSE,warning=FALSE, Plot_Three}
ggplot(data = data,aes(x = alcohol, y = density))+
       geom_point(aes(color =quality))+
       scale_color_continuous(name="quality",low = "white",high = "red")+
       ggtitle("alcohol vs density with different quality")
```

### Description Three

Higher `density` and lower `alcohol`, lower `quality`.

# Reflection

Through this exploratory data analysis, I think feature `alcohol` influence the quality of red wines, however, wine experts give many 5 and 6 score of measure of wine quality, 
maybe just use the data of quality score {3,4} compare to {7,8} will show clearly patterns. I think data visualiztion technique is hard to make decision, it provides too many information, sometimes we just want to know what is the quality of this wine, further study with machine learning could be done to predict the wine quality.